How to Not write Virtual Machines for Dynamic Languages

### 2007-05-27 talk/dyla2007/dyla.tex
some more references, an XXX, a reformulation

In section \ref{sect:approaches} we will explore the way VMs are typically
implemented in C and on top of OO VMs and some of the problems of these
approaches. In \ref{section sect:metaprogramming} we will describe our proposed
meta-programming approach.

@@

@@

These limitations of a C-based implementation often lead to more
implementations of the language. For example in addition to the main C
implementation of Python (``CPython'') there are a number of additional
implementations that want to fix some of the limitations of CPython. One of
them is ``Stackless Python'' which is a fork of CPython that adds micro-threading
capabilities to Python. One of the reasons for not incorporating it back into
CPython was that it was felt that they would make the implementation too
complex. Another implementation of the Python language is the Psyco project
which adds a JIT-compiler to CPython. As all other re-implementations, Psyco
contains the Python semantics and needs to be kept synchronized with CPython
manually.

@@

As a language becomes more popular, yet another reason for different
implementations emerges: People want to use it together with Java and C\#
libraries and tools. Although it is in principle possible to do that by
bridging the C-VM with the respective VMs of these languages (XXX examples)
this is an incomplete solution since it prevents many use-cases.
Therefore in the end implementations of the dynamic language on top of the
object-oriented VMs are created to properly integrate with the complete
environment.

@@

Although it is able to generate, among other things, a complete custom
VM for C-like environments, we found that the required effort that must
be put into the translation toolchain was still much lower than that of
writing a good-quality OO VM.  A reason seems to be that we could design
our translation toolchain specifically for our needs, i.e. a language
implementer's needs, instead of for general-purpose usage.  Of course,
the translation toolchain, once written, can also be reused to implement
other languages, and possibly tailored on a case-by-case basis to fit
each implementer's need.  The process is incremental: we can add more
features as needed instead of starting from a maximal up-front design,
and gradually improve the quality of the tools, the garbage collectors,
the various optimizations, etc.

>>>>

In section \ref{sect:approaches} we will explore the way VMs are typically
implemented in C and on top of OO VMs and some of the problems of these
approaches, using various Python implementations as the main example. In
\ref{section sect:metaprogramming} we will describe our proposed
meta-programming approach.

@@

XXX the next paragraph is also the start of the introduction right now!

@@

These limitations of a C-based implementation often lead to more implementations
of the language. For example in addition to the main C implementation of Python
(``CPython'' \cite{cpy251}) there are a number of additional implementations
that want to fix some of the limitations of CPython. One of them is ``Stackless
Python'' \cite{stackless} which is a fork of CPython that adds micro-threading
capabilities to Python. One of the reasons for not incorporating it back into
CPython was that it was felt that they would make the implementation too
complex. Another implementation of the Python language is the Psyco \cite{Psyco}
project which adds a JIT-compiler to CPython. As all other re-implementations,
Psyco contains the Python semantics and needs to be kept synchronized with
CPython manually.

@@

As a language becomes more popular, yet another reason for different
implementations emerges: People want to use it together with Java and C\#
libraries and tools. Although it is in principle possible to do that by
bridging the C-VM with the respective VMs of these languages (for example the
``PythonNet'' project \cite{pythonnet} that brigdes CPython with a .NET VM)
this is an incomplete solution since it prevents many use-cases.
Therefore in the end implementations of the dynamic language on top of the
object-oriented VMs are created to properly integrate with the complete
environment.

@@

Although it is able to generate, among other things, a complete custom
VM for C-like environments, we found that the required effort that must
be put into the translation toolchain was still much lower than that of
writing a good-quality OO VM.  A reason seems to be that we could design
our translation toolchain specifically for our needs, i.e. a language
implementer's needs, instead of for general-purpose usage.  Of course,
the translation toolchain, once written, can also be reused to implement
other languages, and possibly tailored on a case-by-case basis to fit
the specific need of a language.  The process is incremental: we can add more
features as needed instead of starting from a maximal up-front design,
and gradually improve the quality of the tools, the garbage collectors,
the various optimizations, etc.
